Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2013-01-14 chi2013-webbox.tex
edits thru methodology

In summary, we found in our PS.2, that while data records ranged in complexity, no single source subsumed any others; all contained unique properties.  Moreover, the most common types of heterogeneity exhibited were terminological and value-related, which occurred vastly more than structural or modeling issues.

@@

The primary goal of DataPalette was to empower end-users so that they are able to mix arbitrary, structured data feeds on-the-fly while solving particular information task(s) more effectively.  Unlike mashup makers, such as Yahoo Pipes!, the aim was to facilitate such integration without the need for a separate step involving programming, scraping, shaping, mapping or integration.

@@

The two pre-studies above provided valuable insights towards this goal, which, in turn, informed our design process. PS.1 confirmed the need for data integration, by showing that people preferred to rely upon multiple sources of information for making important decisions, and identified the kinds of sources most commonly used.  PS.2, meanwhile, revealed that the most common kinds of heterogeneity across data sources were of the terminological or structural varieties. As these were the ``simpler'' kinds of heterogeneity, this gave us optimism that an interaction-based approach would be feasible.

In the following sections, we describe the design process of creating the prototype of DataPalette.  This is followed detailed descriptions of prototype with justifications grounded in the pre-studies.

\subsection{Design Process}
We followed a five-phase spiral model consisting of planning, designing, mocking-up, prototyping, and testing to derive the design of DataPalette.  This process was used because of the high degree of initial uncertainty associated with how we'd achieve effective data mixing by end-users. We approached this design challenge through alternative generation and elimination; alternatives were designed, drawn up and evaluated by colleagues (and in later iterations, non-expert potential users), who critiqued the alternatives. The ease with which designs could be prototyped in HTML5 with modern web frameworks allowed us to perform this develop-test-iterate rapidly through seven iterations of this process.

In order to feel familiar to new users, we started with the look and feel from OS X Finder-style file managers.  In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can freely drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label. Unlike file managers, the basic unit of data was not a file, but an \emph{instance} --- corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data. In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data into discrete instances with properties. This was done at time of import, and could be done for every major Web data feed type (JSON, XML, RDF) without any loss of generality.

\subsubsection{Multi-path selection: link-sliding for heterogeneous sets}

Typically, users are interested in viewing and comparing the \emph{properties} of instances; for example, a user might be keen to examine a product's price, rating, manufacturer and so on.  An instance's properties are displayed beneath it; selecting a property creates a \emph{path selection}, which dereferences all of the instances in a group that contain the selected property and displays their corresponding values.  For example, when a user selects the \emph{manufacturer} property for a product, all products in the same group with a manufacturer property will be dereferenced, causing the manufactures to be displayed alongside their corresponding products.  When such a path selection is applied, the properties displayed are updated to be the set of properties of the \emph{terminating value of the path}, so that this value can be dereferenced further.  For example, once selecting a manufacturer, a user may want to know the product's manufacturer's reputation and selects their ``reputation score'' property, causing the previous path selection to be extended. This process is illustrated in Figure \ref{fig:multipathing}.

@@

When instances cannot be dereferenced by the chosen path selection, an additional path selection can be created. As with the first path selection, it causes all matching instances to be dereferenced.  Each group can carry an arbitrary number of path selections, and each path selection can be extended continually by selecting successive properties of values. This achieves the ability to \emph{link slide} as introduced by Parallax~\cite{parallax}, but across sets of heterogeneous items by supporting multiple parallel dereferenced paths simultaneously.  Multiple selection group-dereferencing results in two key benefits. First, that users can quickly compare all instances that have the same structure, and second, that instances with different property names and be quickly consolidated.

%
%\begin{enumerate}
%\item Clicking on a property dereferences all corresponding values makes it easy to quickly compare all instances that have the same structure.  Although we were concerned that users might be surprised that path selection applied to all objects within each group in parallel, users adapted very quickly and used this functionality to save them time.   
%
%\item Supporting dereferencing along different paths \emph{by structure} allows instances with different property names, and even sub-structures to be quickly consolidated by separate dereferencing, thereby effectively coping with terminological and structural heterogeneity.
%\end{enumerate}
%
\subsubsection{Coreference consolidation: Drag-and-Drop Same-as}

A major challenge with integration from multiple sources is coreference consolidation, or combining representations that co-refer to the same entity into one.  For example, consolidating one's friends from Facebook, Google+, and LinkedIn, one might quickly want to de-duplicate and combine friends that use multiple services so that they each have one representation.  Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, identifiers used by each network were system-specific and inconsistent. Thus, we sought to instead make it user-driven, and easy to perform manually.

@@

To remain consistent with the drag-and-drop interactions used throughout the interface, we introduced the ability to simply drag one instance on top of another to specify that two things should be combined.  This action also caused the \emph{same-as display} in the lower left of the screen to display this new combination relationship.  The display also acted as a mechanism to delete/undo combinations, which reverted representations of instances back to their original, separate forms. An illustration of this interaction is provided in Figure \ref{fig:sameas}. Although general, this can be tedious for groups of items.  Thus, to support more efficient combination, we extended this to allow entire groups to be dropped on other groups.  When this is done, DataPalette attempts to identify matching pairs among items in the dragged and target groups by comparing each pairs' labels.  Only exact matches (modulo white space/underscores) were considered matches and automatically combined.  The rest are added to the target group as distinct elements, and may be explicitly combined by hand.\looseness=-1

@@

\subsubsection{Enumerated-type value consolidation}
PS.2 revealed that enumerated values of properties were often inconsistent, e.g. ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property. To reconcile and consolidate corresponding values from different  data sources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.

@@

To make it easier for people to quickly visually compare aspects of instances, we created a charting tool and integrated a mapping feature.  To make these tools suitable for fast use, e.g. data exploration, we designed these tools to automatically configure display parameters instances or groups of were dropped on them, rather than making users manually specify them. In the current prototype, the DWIM behaviour of the chart automatically sets whether the plot is a histogram representation (counts of values) or numeric value display based upon the types of the dereferenced values being plotted; values that are non-numeric are counted, while numeric values are displayed.    Similarly, the map was made to detect address strings and latitude/longitude pairs, geocoding them where appropriate, and determining the optimal bounding box to display all items at maximum zoom.

\subsubsection{Model-based brushing}
Since it can get confusing having a single instance represented across several groups, we provided \emph{universal brushing} across the interface; hovering over any representation of any instance (an instance block, a map marker or a histogram bar), causes all other visible representations of that instance to lightly glow, so that the user can quickly identify all the places that a single instance is represented.

@@

In order to determine whether the interaction affordances introduced in DataPalette (henceforth DP) allow users to perform serendipitous data mixing on real data feeds by end-users, we devised a within-subjects controlled lab study which we describe next.

@@

We first describe our methodology, starting with the three hypotheses we set out to test:
%These hypothesis pertain to the usability of DP by end-users (h1), whether DP achieves its stated goal of enabling end-users to integrate data (h2), and finally if it has an effect on task performance (h3) - either in how well people perform a task, how quickly they perform it, or if it reduces task effort, as follows:

@@

The participants were seated at a standard OS X desktop equipped with a 22-inch monitor with a standard keyboard and mouse. Each participant was given a maximum of 10 minutes to each of two tasks, but were told that if they finished earlier they could tell us and move on to the next condition.  Prior to A conditions, participants were trained on how to use DP using a five-minute instructional video.

In the control condition (B), participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from web sites pertaining to the candidates they were asked to choose from. Participants were told they were free to use the web sites and spreadsheets however they wished.

% why? why were these tasks selected? can we drop in the phrase ``multi-dimensional subjective choice task'' and refer back to PS.1 to justify the choice of tasks?
We based our two tasks on choosing a university and picking a restaurant for a large social event, because in PS.1 the majority of people stated that they had used multiple websites for these tasks.  We used two different variables for both tasks, where the restaurant was located (Glasgow and Cambridge) and which subject people chose (sport science and history).  For the restaurant task, the cities Glasgow and Cambridge were chosen because they were geographically distant to participants, and would therefore minimise likelihood of prior familiarity with the restaurants or location.  For the University selection task, two university courses were selected at random from the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}, and the six top-ranked universities for each subject were selected as candidates.  
% \subsubsection{Task preparation} In order to prepare the data for our tasks, we identified websites that people commonly used for selecting universities and restaurants (see Table \ref{tab:studyfactors}).  Once we had compiled the list of websites, we select the top most used six for each task. 
We selected six because it enables us to scope our study, we found through two pilot studies that six data sources provided a sufficient challenge for users to explore the data and required them to use DP's main features. We manually collected the key data points from the websites for use in DP and into Excel spreadsheets (which 40\% of the participants in PS.1 said that had previously used to compare data).

@@

\subsection{Data Collection and Analysis}
Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.  Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn't influence the study.   Explicitly both the facilitator and observer where trained in identifying processes that people used during a task, so that they could thematically categorise them.  Initially, we identified the following themes: organisation of data, eliminating candidates, co-referencing, and visualisations. The studies were run over 4 days.  At the end of each day the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.  During the lab study, we recorded the audio and the participant's actions on screen. We asked the participants to follow a \emph{think-aloud} protocol as they worked, so that we could understand the reasoning behind their actions.  At the end of the study, the participant completed a short exit survey.

After the studies were complete, we generated transcripts of comments made by participants during the think-aloud protocols. At the same time, we populated a spreadsheet summarising quantitative metrics pertaining to how much various features of the tool were used.  After the transcriptions were complete, we categorised people's comments made during the study, suggests in their exit survey, and clustered them to common themes.

@@

\subsection{Participant Recruitment}
We recruited participants through adverts posted near the University campus and e-mails to student mailing lists.  We screened participants to be at least 18 years of age, but did not filter on any other criteria.  We took the first 20 that signed up and that successfully specified time slots that were available.   Participants were offered a small gift certificate from an online retailer for their time.

@@

We received 26 responses to our call, from which we selected the first 20 (ten of whom were female). Seven were non-students, comprised of university staff and alumni; the remainder were students, 8 studying computer science, and 5 studying other subjects.  Due to the large population of international students and staff at our university, we ended up with a large sample (12) of non-native English speakers; however, since all had passed requisite proficiency tests, we did not think this would substantially impact results.  On average, each study took 40 minutes to complete.  %The information in Table \ref{tbl:data} summarises the captured metrics for each participant.

@@

Our study had several limitations.  First, the small sample size of our study (20 participants) meant that we could not confirm some of the effects observed with statistical significance.  Second, giving participants a time limit of 10 minutes on trials may have affected the strategies they used to make their choices.  For this reason, in our follow-up study, we will give some participants unlimited time on tasks of similar complexity.  A second departure from a realistic task was the way in which we pre-determined the data sources for participants to use.  We did this in order to focus participants' efforts on looking at and working with their choices, rather than retrieving it.  However, as supporting effective retrieval and serendipitous discovery will be important for DP to be useful in practice, we plan to look at supporting these stages in the future. 
% \subsection{Integration Limitations}

% In the A condition, where participants used DataPalette to complete their tasks, they used it to collate the data on-screen, and compare multiple values directly. About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it. Those who tried to use it were largely successful, and were able to therefore use the global brush-highlighting to simultaneously compare different statistics. Most users instantly opened up all data sources into individual boxes, in order to determine which sources contained specific statistics. The result of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.

%% \subsection{Better Supporting the Task}

%% - are people more thorough
%% - do they evaluate more sources

@@

As said in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for fully-automatic, complete approaches to data integration have casued simpler, manual approaches to be overlooked.  Based upon the results of our investigation, we see a number of benefits to interactive, lightweight approaches over automatic and bespoke methods, at least for several classes of common, information-gathering for decision-making tasks.

We believe that the key affordances for the suitability of such tools are flexibility, simplicity, and ease of use.  Just as work in the ``user-subjective approach'' to PIM (e.g., \cite{user-subjective}) showed that personal, user-specified attributes were ultimately the most useful, granting users the flexibility to combine or separate information items as needed allowed the tool to support the varied strategies participants employed to compare and examine their options. Second, the low cost of use  (in terms of time and effort required) meant that integration could be done (and undone) repeatedly within the ``inner loop'' of their exploratory decision-making process.  The importance of this was that it was the process itself that determined, incrementally, what information was ultimately needed, including the sources that were consulted. Therefore, the use of any integration approach that required sources to be determined \emph{a priori}, or that required a time or effort-intensive integration step would not have been feasible.

Based on these findings, we plan to continue investigating lightweight, \emph{in-situ} approaches to supporting ad-hoc data integration during the course of sensemaking and decision-making activities.  In particular, we are examining gestural input methods for the purposes of supporting more fluid exploration and mixing.  Based upon feedback from reviewers, we have also begun considering semi-automatic and crowd-powered methods to help users quickly resolve challenging reconciliation problems they might encounter.

>>>>

In summary, we found in our PS.2, that while data records ranged in complexity, no single source subsumed any others; all contained unique properties.  Moreover, the most common types of heterogeneity exhibited were terminological and value-related, which occurred vastly more often than structural or modeling issues.

@@

The primary goal of DataPalette (henceforth DP) was to empower end-users so that they were able to mix arbitrary, structured data feeds on-the-fly, while solving particular information task(s) more effectively.  Unlike mashup makers, such as Yahoo Pipes!, the aim was to facilitate such integration without the need for a separate step involving programming, scraping, shaping, mapping or integration.

@@

The two pre-studies above provided valuable insights towards this goal, which, in turn, informed our design process. PS.1 confirmed the need for data integration, by showing that people preferred to rely upon multiple sources of information for making important decisions; PS.1 also identified the kinds of sources most commonly used.  PS.2, meanwhile, revealed that the most common kinds of heterogeneity across data sources were of the terminological or structural varieties. As these were the ``simpler'' kinds of heterogeneity, this gave us optimism that an interaction-based approach would be feasible.

In the following sections, we describe the design process of creating the prototype of DP.  This is followed by detailed descriptions of the prototype, with justifications grounded in the pre-studies.

\subsection{Design Process}
We followed a five-phase spiral model, consisting of planning, designing, mocking-up, prototyping, and testing, to derive the final design of DP.  This multi-phase process was used, due to the high degree of initial uncertainty associated with how users would achieve effective data mixing. We approached this design challenge through alternative generation and elimination; alternatives were designed, drawn up and evaluated by colleagues (and in later iterations, non-expert potential users), who critiqued the alternatives. The ease with which designs could be prototyped in HTML5 with modern web frameworks allowed us to perform this process of develop-test-iterate rapidly through seven iterations.

In order to allow new users to be comfortable with our interface, we adopted the look and feel from OS X Finder-style file managers.  In the DP workspace, shown in Figure \ref{fig:workspace}, the user would be able to freely drag and drop entities from data sources (on the left) into windows that they have created, resized, arranged and labeled. Unlike file managers, the basic unit of data was not a file, but an \emph{instance} --- corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data. In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data into discrete instances with properties. This was done at time of import, and could be done for every major Web data feed type (JSON, XML, RDF) without any loss of generality.

\subsubsection{Multi-path selection: link-sliding for heterogeneous sets}
Typically, users are interested in viewing and comparing the \emph{properties} of instances; for example, a user might be keen to examine a product's price, rating, manufacturer and so on.  To support such comparisons, an instance's properties would be displayed beneath it; selecting a property would create a \emph{path selection}, which would dereference all of the instances in a group that contained the selected property and displayed their corresponding values.  For example, when a user selected the \emph{manufacturer} property for a product, all products in the same group with a manufacturer property would be dereferenced, causing the manufacturers to be displayed alongside their corresponding products.  When such a path selection was applied, the properties displayed were updated to be the set of properties of the \emph{terminating value of the path}, so that this value could be dereferenced further.  Having selected a manufacturer, a user might want to be informed of its reputation.  By selecting the ``reputation score'' property, the previous path selection would then be extended. This process is illustrated in Figure \ref{fig:multipathing}.

@@

When instances could not be dereferenced by the chosen path selection, an additional path selection could be created. As with the first path selection, this would cause all matching instances to be dereferenced.  Each group would be able to carry an arbitrary number of path selections, and each path selection could be extended continually by selecting successive properties of values. This would achieve the ability to \emph{link slide} multiple instances simultaneously, as introduced by Parallax~\cite{parallax}, extended to support sets of heterogeneous items through multiple parallel simultaneous paths.  Multiple-path group-dereferencing would result in two key benefits. First, users would be able to quickly compare all instances that have the same structure, and second, instances with different property names would be easily consolidated.

%
%\begin{enumerate}
%\item Clicking on a property dereferences all corresponding values makes it easy to quickly compare all instances that have the same structure.  Although we were concerned that users might be surprised that path selection applied to all objects within each group in parallel, users adapted very quickly and used this functionality to save them time.   
%
%\item Supporting dereferencing along different paths \emph{by structure} allows instances with different property names, and even sub-structures to be quickly consolidated by separate dereferencing, thereby effectively coping with terminological and structural heterogeneity.
%\end{enumerate}
%
\subsubsection{Coreference consolidation: Drag-and-Drop Same-as}
A major challenge with integration from multiple sources is coreference consolidation, or combining into one, representations that co-refer to the same entity.  For example, when consolidating one's friends from Facebook, Google+, and LinkedIn, one might wish to eliminate duplication and combine friends' profiles across multiple services.  Although we considered ways to do this automatically, PS.2 revealed no viable method to reliably identify the co-referring instances; due to a lack of standard URIs across social networks, identifiers used by each network were system-specific and inconsistent. Thus, we sought to instead make it user-driven, and easy to perform manually.

@@

To remain consistent with the drag-and-drop interactions used throughout the interface, we introduced the ability to simply drag one instance directly on top of another, to specify that these two items should be combined.  This action also caused the \emph{same-as display} in the lower left of the screen to indicate this new combined relationship.  The display also acted as a mechanism to delete/undo combinations, which reverted representations of instances back to their original, separate forms. An illustration of this interaction is provided in Figure \ref{fig:sameas}.  Since individually dragging and dropping items could be tedious for large groups of items, we extended this to allow entire groups to be dropped on to other groups.  DP would then identify matching pairs among items in the dragged and target groups by comparing the labels of items in each group, repsectively.  Only exact matches (excepting white space and underscores) were considered to correspond to label matches, and these were automatically combined.  The rest were added to the target group as distinct elements, which could be manually combined by hand.\looseness=-1

@@

\subsubsection{Enumerated-type value consolidation}
PS.2 revealed that enumerated values of properties were often inconsistent, e.g., ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property. To reconcile and consolidate corresponding values from different  data sources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.

@@

To make it easier for people to visually compare aspects of instances, we created a charting tool and mapping feature.  To ensure that these tools were suitable for rapid utilisation, (e.g., data exploration), they were designed to automatically configure display parameters based on the instances or groups that were dropped onto them. In the current prototype, this ``Do What I Mean'' behaviour automatically would set whether the plot was a histogram representation (counts of values) or a numeric value display, based upon the types of the dereferenced values being plotted.  Hence, non-numeric values would be histogrammed, while numeric values were displayed as a bar or line chart.  Similarly, the map was made to detect address strings and latitude/longitude pairs, geocoding them where appropriate, and determining the optimal bounding box to display all items at maximum zoom.

\subsubsection{Model-based brushing}
To avoid the confusion of having a single instance represented across several groups, we provided \emph{universal brushing} across the interface.  Hovering over any representation of any instance (an instance block, a map marker or a histogram bar), would cause all other visible representations of that instance to lightly glow, enabling the user to instantly identify all the places that a single instance was represented.

@@

In order to determine whether or not the interaction affordances introduced in DP allowed users to perform serendipitous data mixing on real data sources, we devised a within-subjects controlled lab study, which we describe next.

@@

We began with three hypotheses pertaining to people's ability to use DP, and its impact on task performance:
%These hypothesis pertain to the usability of DP by end-users (h1), whether DP achieves its stated goal of enabling end-users to integrate data (h2), and finally if it has an effect on task performance (h3) - either in how well people perform a task, how quickly they perform it, or if it reduces task effort, as follows:

@@

The participants were seated at a standard OS X desktop equipped with a 22-inch monitor with a standard keyboard and mouse. Each participant was given a maximum of 10 minutes to complete each of two tasks, but were told that if they finished early, they could inform us and move on to the next condition.  Prior to the A condition trials, participants were trained on how to use DP using a five-minute instructional video.

In the control (B) condition trials, participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from web sites pertaining to the candidates they were asked to choose from. Participants were told they were free to use the web sites and spreadsheets however they wished.

% why? why were these tasks selected? can we drop in the phrase ``multi-dimensional subjective choice task'' and refer back to PS.1 to justify the choice of tasks?

We established that our two tasks would entail choosing a university and selecting a restaurant for a large social event.  These tasks were selected because in PS.1, the majority of people stated that they had used multiple websites for such tasks.  We introduced a variable for each task.  For the University selection task, we varied the course for which to select a University, between history and sports science, which we had selected randomly from the course list of the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}.  For the restaurant selection task, the variable selected where the restaurants were located; for this experiment, the cities Glasgow and Cambridge were chosen because they were geographically distant from participants, and would therefore minimise prior familiarity.

For each of the two tasks, the final candidates from which participants made their selection were pre-selected by us.  For the university task, the six top-ranked universities for each subject according to The Complete University Guide were selected as candidates.  We selected six in order to limit the difficulty of the task given the fixd time allocated for each trial.  We manually collected the key statistics about each university from each of the data sources, and collated them into Excel spreadsheets.  % In addition, for each subject, we provided a hypothetical student's A/AS-level scores and personal preferences pertaining to distance from home, %%

For the restaurant task, we selected six restaurants for each location randomly from Yelp's list of popular eateries for each of the two cities.  In addition, for each of the two cities, we generated 12 social network profiles for hypothetical friends who lived in either city, to accompany the participant to the restaraunt.  Each friend's profile expressed their personal preferences for favourite cuisine types, which could be taken into consideration by participants during the selection process.

% We established that our two tasks would entail choosing a university and selecting a restaurant for a large social event.  These tasks were selected because in PS.1, the majority of people stated that they had used multiple websites for such tasks.  We used two different variables for each task.  where the restaurant was located (Glasgow and Cambridge) and which subject people chose (sport science and history).  For the restaurant task, the cities Glasgow and Cambridge were chosen because they were geographically distant to participants, and would therefore minimise likelihood of prior familiarity with the restaurants or location.  For the University selection task, two university courses were selected at random from the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}, and the six top-ranked universities for each subject were selected as candidates.  We selected six because it enables us to scope our study, we found through two pilot studies that six data sources provided a sufficient challenge for users to explore the data and required them to use DP's main features. We manually collected the key data points from the websites for use in DP and into Excel spreadsheets (which 40\% of the participants in PS.1 said that had previously used to compare data).

@@

\subsection{Data Collection and Analysis}
Each study was overseen by a facilitator and an observer: the role of the facilitator was to explain the study's protocols to the participant and to answer any questions; the role of the observer was to monitor the study and record observations.  Both the facilitator and observer were trained on the purpose of the study, their respective roles, and how they could or couldn't influence the study.   Explicitly, the facilitator and the observer were trained to identify processes that participants might use during a task, so that they could thematically categorise them.  Initially, we identified the following themes: organisation of data, eliminating candidates, co-referencing, and visualisations. The studies were conducted over a period of 4 days.  At the end of each day, the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.  During the lab study, we recorded the audio and the participant's actions on screen. We asked the participants to follow a \emph{think-aloud} protocol as they worked, so that we could understand the reasoning behind their actions.  At the end of the study, the participant completed a short exit survey.

After the studies were complete, we generated transcripts of comments made by participants during the think-aloud protocols. At the same time, we created a spreadsheet summarising quantitative metrics pertaining to how much of the various features of DP were used.  After the transcriptions were completed, we categorised comments made by the participants during the study, together with suggestions from their exit survey, and clustered them to common themes.

@@

\subsection{Participant Recruitment}

We recruited participants through adverts posted near the Southampton University campus and via e-mails to student mailing lists.  We screened participants to ensure they were at least 18 years of age, but did not filter candidates based on any other criteria.  We received 26 responses to our call, from which we accepted the first 20 applicants (10 of whom were female). Seven were non-students, comprised of university staff and alumni; the remainder were students, 8 studying computer science, and 5 studying other subjects.  Due to the large population of international students and staff at our university, we ended up with a large sample (12) of non-native English speakers; however, since all had passed requisite proficiency tests, we did not think this would substantially impact results.  On average, each study took 40 minutes to complete.  Participants were offered a small gift certificate from an online retailer for their time.

@@

% We received 26 responses to our call, from which we selected the first 20 (ten of whom were female). Seven were non-students, comprised of university staff and alumni; the remainder were students, 8 studying computer science, and 5 studying other subjects.  Due to the large population of international students and staff at our university, we ended up with a large sample (12) of non-native English speakers; however, since all had passed requisite proficiency tests, we did not think this would substantially impact results.  On average, each study took 40 minutes to complete.  %The information in Table \ref{tbl:data} summarises the captured metrics for each participant.

@@

Our study had several limitations.  First, the small sample size of our study (20 participants) meant that we could not confirm some of the effects observed with statistical significance.  Second, giving participants a time limit of 10 minutes on trials may have affected the strategies they used to make their choices.  For this reason, in our follow-up study, we will give some participants unlimited time on tasks of similar complexity.  A second departure from a realistic task was the way in which we pre-determined the data sources for participants to use.  We did this in order to focus participants' efforts on looking at and working with their choices, rather than retrieving them.  However, as supporting effective retrieval and serendipitous discovery will be important for DP to be useful in practice, we plan to look at supporting these stages in the future.

% \subsection{Integration Limitations}
% In the A condition, where participants used DataPalette to complete their tasks, they used it to collate the data on-screen, and compare multiple values directly. About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it. Those who tried to use it were largely successful, and were able to therefore use the global brush-highlighting to simultaneously compare different statistics. Most users instantly opened up all data sources into individual boxes, in order to determine which sources contained specific statistics. The result of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.

%% \subsection{Better Supporting the Task}
%% - are people more thorough
%% - do they evaluate more sources

@@

As expressed in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for fully-automatic, complete approaches to data integration have caused simpler, manual approaches to be overlooked.  Based upon the results of our investigation, we have seen a number of benefits to interactive, lightweight approaches over automatic and bespoke methods, at least for a frequently-occurring class of information gathering for decision-making tasks.

In particular, the results suggest that the key affordances for the suitability of such tools are flexibility, simplicity, and ease of use.  Just as work in the ``user-subjective approach'' to PIM (e.g., \cite{user-subjective}) showed that personal, user-specified attributes were ultimately the most useful, granting users the flexibility to combine or separate information items as needed allowed the tool to support the varied strategies participants employed to compare and examine their options. Second, the low cost of use  (in terms of time and effort required) meant that integration could be done (and undone) as appropriate throughout the ``inner loop'' of users' exploratory processes.  The reason this was significant was that this process itself determined, incrementally, the information and sources that were ultimately considered for the final decision. Therefore, any integration approach that required sources to be specified \emph{a priori}, or that was time or effort-intensive at each integration step, was simply not appropriate for this context.

Due to the promising findings of this study, we plan to continue investigating lightweight, \emph{in-situ} approaches to supporting ad-hoc data integration during the course of sensemaking and decision-making activities.  In particular, we are examining gestural input methods for the purposes of supporting more fluid exploration and mixing.  On the basis of feedback from reviewers, we have also begun considering semi-automatic and crowd-powered methods to help users quickly resolve challenging reconciliation problems they might encounter.
